---
title: "R Notebook"
output: html_notebook
---
In this problem, you will use simulation to evaluate (by Monte Carlo) the expected misclassification error rate given a particular generating model.

Let yi be equally divided between classes 0 and 1, and let xi belongs to R^10 be normally distributed.

 Given yi = 0, x ~ N10(0, I10).
 Given yi = 1, x ~ N10(mu, I10) with mu = (1,1,1,1,1,0,0,0,0,0)

The  notation just means its a ten-dimensional Gaussian distribution, you can use the mvrnorm function in the MASS package to help generate the data.

Now, we would like to know the expected test error rate if we fit an SVM to a sample of 50 random training points from class 1 and 50 more from class 0.  We can calculate this to high precision by 
1) generating a random training sample to train on,
2) evaluating the number of mistakes we make on a large test set, 
and then 3) repeating (1-2) many times and averaging the error rate for each trial. 

```{r}
library(MASS)
library(e1071)

set.seed(1001)
count <- 1000
err_rate <- integer()

for(i in 1:count)
{
  x <- matrix(rnorm(100 * 10), ncol = 10)
  y <- c(rep(0,50), rep(1,50))
  x[y == 1, 1:5] <- x[y == 1, 1:5] + 1
  
  dat <- data.frame(x = x, y = as.factor(y))
  
  svm_fit <- svm(y~., data = dat, kernel = "radial", cost = 1)
  
  xtest <- matrix(rnorm(100*10), ncol = 10)
  ytest <- sample(c(0,1), 100, rep = TRUE)
  xtest[ytest == 1, 1:5] <- xtest[ytest==1, 1:5] + 1
  testdat <- data.frame(x = xtest, y = as.factor(ytest))
  
  ypred <- predict(svm_fit, newdata = testdat)
  result <- table(predict = ypred, truth = testdat$y)
  
  err_rate[i] <- 1 - (result[1] + result[4] )/ 100
}
mean(err_rate)
```


```{r}
library(MASS)

set.seed(1001)
count <- 1000
err_rate <- integer()

for(i in 1:count)
{
  x <- matrix(rnorm(100 * 10), ncol = 10)
  y <- c(rep(0,50), rep(1,50))
  x[y == 1, 1:5] <- x[y == 1, 1:5] + 1
  
  dat <- data.frame(x = x, y = as.factor(y))
  
  logit_fit <- glm(y~., data = dat, family = "binomial")

  xtest <- matrix(rnorm(100*10), ncol = 10)
  ytest <- sample(c(0,1), 100, rep = TRUE)
  xtest[ytest == 1, 1:5] <- xtest[ytest==1, 1:5] + 1
  testdat <- data.frame(x = xtest, y = as.factor(ytest))
  
  ypred <- predict(logit_fit, newdata = testdat, type = "response")
  ypredType <- ifelse(ypred > 0.5, 1, 0)
  
  result <- table(predict = ypredType, truth = testdat$y)
  
  err_rate[i] <- 1 - (result[1] + result[4] )/ 100
}
mean(err_rate)

```

