---
title: "ISLR - Learning Classification using the Stock Market Data"
output: html_notebook
---

1) Running Logistic Regression models on the Stock Market dataset

```{r}
library(ISLR)
names(Smarket)
summary(Smarket)
```

```{r}
pairs(Smarket, col = Smarket$Direction)
```

- Running a logistic regression model

```{r}
glm_fit <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = Smarket, family = binomial)
summary(glm_fit)
glm_probs <- predict(glm_fit, type = "response")
glm_probs[1:5]
glm_pred <- ifelse(glm_probs>0.5,"Up","Down")
```


Checking our performance

```{r}
#Getting the classification confusion matrix
table(glm_pred, Smarket$Direction)

#Getting our classification performance
#Correct classifications/Total number of observations
mean(glm_pred==Smarket$Direction)
```

- Splitting into train-test
```{r}
train <- Smarket$Year < 2005
glm_fit_train <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = Smarket, family = binomial, subset = train)
glm_probs_test <- predict(glm_fit_train, newdata = Smarket[!train,], type = "response")
glm_pred_test <- ifelse(glm_probs_test>0.5, "Up", "Down")
```

Checking performance on test data
```{r}
Smarket_Direction_2005 <- Smarket$Direction[!train]
table(glm_pred_test, Smarket_Direction_2005)

#Correct classification accuracy for test
mean(glm_pred_test == Smarket_Direction_2005)
```

- Classification using only lag1 and lag2 terms
```{r}
glm_fit_train2 <- glm(Direction ~ Lag1 + Lag2, data = Smarket, family = binomial, subset = train)
glm_probs_test2 <- predict(glm_fit_train2, newdata = Smarket[!train,], type = "response")
glm_pred_test2 <- ifelse(glm_probs_test2>0.5, "Up", "Down")
```

Checking performance of updated classification codes
```{r}
table(glm_pred_test2, Smarket_Direction_2005)
mean(glm_pred_test2 == Smarket_Direction_2005)

```



----------------------------------------------------------------------

2) Running Linear Discriminant Analysis on the Stock Market dataset

```{r}
library(MASS)
rm(list = ls())
```


Fitting a LDA model
```{r}
lda_fit <- lda(Direction ~ Lag1 + Lag2, data = Smarket, subset = Year < 2005)
lda_fit
plot(lda_fit)
```


Checking what the performance is on test data (year = 2005)

```{r}
Smarket_2005 <- subset(Smarket, Year == 2005)
lda_pred = predict(lda_fit, Smarket_2005)
data.frame(lda_pred)[1:5,]

table(lda_pred$class,Smarket_2005$Direction)
mean(lda_pred$class==Smarket_2005$Direction)

```


---------------------------------------------------------------------

3) K-Nearest Neighbours classification

```{r}
library(class)
?knn

Xlag <- cbind(Smarket$Lag1, Smarket$Lag2)
train <- Smarket$Year < 2005

knn_pred <- knn(Xlag[train,], Xlag[!train,], Smarket$Direction[train], k=1)

table(knn_pred, Smarket$Direction[!train])
mean(knn_pred == Smarket$Direction[!train])

```



## K-Nearest Neighbors
library(class)
?knn
attach(Smarket)
Xlag=cbind(Lag1,Lag2)
train=Year<2005
knn.pred=knn(Xlag[train,],Xlag[!train,],Direction[train],k=1)
table(knn.pred,Direction[!train])
mean(knn.pred==Direction[!train])
